Visual inference for networks

### 2016-03-31 PaperThoughts.Rnw
fixed small friend lineup so that singletons appear.

\item \textbf{Moving from $x(t_{m-1})$ to $x(t_m)$} - First define $x(t_1)$ as the first network observed and $x(t_M)$ as the last network observed. Networks in between are written $x(t_{m})$ for $m = 2, \dots, M-1$. The observations $x(t_1), x(t_2), \dots, x(t_M)$ are embedded in the continuous-time Markov chain $\{X(t)|t_1 \leq t \leq t_M\}$. Moving from the observation $x(t_{m-1})$ to the observation $x(t_m)$ requires $R$ changes or steps in the Markov process $X(t)$ where $R = \sum_{i,j \in \mathcal{N}} |x_{ij}(t_m) - x_{ij}(t_{m-1})|$. Note that the observed network change from $x(t_{m-1})$ to $x_(t_m)$ is conditionally independent of all other network observations given $x(t_{m-1})$. Denote the $R$ timepoints in the the Markov process that correspond to the unobserved changes in the network as $T_1, T_2, \dots, T_{R-1}$ and define $T_0 = t_{m-1}$ and $T_R \leq t_{m} < T_{R+1}$. For each timepoint $T_r$, for $r = 1, \dots R$, there is a single actor, denoted $I_r$ that gets an opportunity to change at this timepoint. This actor changes one of its ties to one node, $J_r$. If there is no change, $I_r = J_r$. Formally, the pair $(I_r,J_r)$ is the only $(i,j)$ pair for which $x_{ij}(T_{r-1}) \neq x_{ij}(T_r)$ if such a pair exists, and $(I_r, J_r) = (I_r, I_r)$ otherwise. The triplet $(T_r, I_r, J_r)$ forms a stochastic process for $r = 1, \dots, R$ that, given the prior observation, $x(t_{m-1})$ for $m = 2, \dots, M$, completely determines $x(t)$ for $t_{m-1} < t < t_m$. 

>>>>

\item \textbf{Moving from $x(t_{m-1})$ to $x(t_m)$} - First define $x(t_1)$ as the first network observed and $x(t_M)$ as the last network observed. Networks in between are written $x(t_{m})$ for $m = 2, \dots, M-1$. The observations $x(t_1), x(t_2), \dots, x(t_M)$ are embedded in the continuous-time Markov chain $\{X(t)|t_1 \leq t \leq t_M\}$. Moving from the observation $x(t_{m-1})$ to the observation $x(t_m)$ requires $R$ changes or steps in the Markov process $X(t)$ where $R = \sum_{i,j \in \mathcal{N}} |x_{ij}(t_m) - x_{ij}(t_{m-1})|$. Note that the observed network change from $x(t_{m-1})$ to $x_(t_m)$ is conditionally independent of all other network observations given $x(t_{m-1})$. Denote the $R$ timepoints in the the Markov process that correspond to the unobserved changes in the network as $T_1, T_2, \dots, T_{R}$ and define $T_0 = t_{m-1}$ and $T_R \leq t_{m} < T_{R+1}$. For each timepoint $T_r$, for $r = 1, \dots R$, there is a single actor, denoted $I_r$ that gets an opportunity to change at this timepoint. This actor changes one of its ties to one node, $J_r$. If there is no change, $I_r = J_r$. Formally, the pair $(I_r,J_r)$ is the only $(i,j)$ pair for which $x_{ij}(T_{r-1}) \neq x_{ij}(T_r)$ if such a pair exists, and $(I_r, J_r) = (I_r, I_r)$ otherwise. The triplet $(T_r, I_r, J_r)$ forms a stochastic process for $r = 1, \dots, R$ that, given the prior observation, $x(t_{m-1})$ for $m = 2, \dots, M$, completely determines $x(t)$ for $t_{m-1} < t < t_m$. 

### 2016-03-31 paper/network-inference.Rnw
now the comments get comments
fixed small friend lineup so that singletons appear.

\par Social networks are generally not static: they evolve over time. This evolution is influenced by both the actors' behaviors and the network structure. To deal with this continuous evolution, 
stochastic actor-oriented models (SAOMs) were introduced by Snijders in~\cite{Snijders:1996}. 
In a SAOM each actor is represented by a node. Each node has its own objective function, $f_i(\beta, x)$ where $i$ is the node index, and $\beta$ is a vector of $p$ coefficients for the $p$ effects in a model. The full form of the objective function is 
\begin{equation}\label{eq:SAOM}
f_i(\beta, x) = \sum_k \beta_k s_{ik}(x),
\end{equation}
where $s_{ik}(x)$ \hh{is a structural function ... ??? XXX do these functions have a specific name?} for network state $x$, effect $k = 1, ..., p$ and node $i$.

@@

\st{Alternatively, the actor can do nothing. \hh{XXX how is that decided? somehow the probabilities $p_ij$, $j=1, ..., ?$ have to be evaluated and a decision in favor of acting or not is made. from the description it's not clear yet how that happens.} These single edge change moments are unobserved. They are only observed in aggregate when the network is observed at discrete time points. So, multiple actors may change (multiple of) their ties between wave $m-1$ and wave $m$.   
\par

>>>>

\par Social networks are generally not static: they evolve over time. This evolution is influenced by both the actors' behaviors and the network structure. To deal with this continuous evolution, 
stochastic actor-oriented models (SAOMs) were introduced by Snijders in~\cite{Snijders:1996}. A SAOM is a discrete process embedded in a continuous-time Markov process of network evolution. In the Markov chain, each step consists of adding or deleting one tie $(i,j)$. So, the change from the first observation of a network, call it $x(t_1)$, to the next observation, $x(t_2)$ is modeled as a series of $R$ changes, where $R = \sum_{i,j} |x_{ij}(t_2) - x_{ij}(t_1)|$. These individual changes are unobserved, and during the estimation of parameters, they serve as augmenting data. Denote the $R$ timepoints in the the Markov process that correspond to the unobserved changes in the network as $T_1, T_2, \dots, T_{R}$ and define $T_0 = t_{1}$ and $T_R \leq t_{2} < T_{R+1}$. For each timepoint $T_r$, for $r = 1, \dots R$, there is a single actor, denoted $I_r$ that gets an opportunity to change at this timepoint. This actor changes one of its ties to one node, $J_r$. If there is no change, $I_r = J_r$. The waiting times, $T_r - T_{r-1}$ for $r = 1, \dots, R$ have independent exponential distributions with expected value $(\lambda(\alpha, x^{(r-1)}))^{-1}$. The function $\lambda(\alpha, x^{(r-1)})$ is the rate at which the network changes. Here, $x^{(r)} \equiv x(T_r)$ for $r = 2, \dots, R$. It is the sum of the rates at which each node makes a change in one of its ties, $\lambda_i(\alpha, x^{(r-1)})$. Note that x^{(r)}
In a SAOM each actor is represented by a node. Each node has its own objective function, $f_i(\beta, x)$ where $i$ is the node index, and $\beta$ is a vector of $p$ coefficients for the $p$ effects in a model. The full form of the objective function is 
\begin{equation}\label{eq:SAOM}
f_i(\beta, x) = \sum_k \beta_k s_{ik}(x),
\end{equation}
where $s_{ik}(x)$ \hh{is a structural function ... ??? XXX do these functions have a specific name?} for network state $x$, effect $k = 1, ..., p$ and node $i$.

@@

\st{Alternatively, the actor can do nothing. \hh{XXX how is that decided? somehow the probabilities $p_ij$, $j=1, ..., ?$ have to be aggregated and a decision in favor of acting or not must be made. from the description it's not clear yet how that happens.} These single edge change moments are unobserved. They are only observed in aggregate when the network is observed at discrete time points. So, multiple actors may change (multiple of) their ties between wave $m-1$ and wave $m$.   
\par
